---
title: "Aspen Meadow"
output: html_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# why can I not use source("scripts/setup.R")?

source("../../scripts/setup.R")
```

# Ground Cover

Let's read in the Aspen Meadows ground cover data for Fall 2024.

```{r}
# reading in the ground cover data for Fall 2024
aspen_cover_2024 <- 
  read_excel("../../data/clean/aspen_meadow.xlsx", sheet = "F24 % Ground Cover")

view(aspen_cover_2024)
```

Next, let's select the information we need to create a plot comparing the ground cover across transects.

```{r}
# select only transect and bare ground types
aspen_cover_types_2024 <-
  aspen_cover_2024 %>%
  select(1, 3:9)

view(aspen_cover_types_2024)
```

To more efficiently work with this graph, let's "tidy" the data.

```{r}
aspen_cover_types_2024_tidy <- aspen_cover_types_2024 %>%
  pivot_longer(cols = !Transect, names_to = "Cover_Type", 
               values_to = "Percentage")

view(aspen_cover_types_2024_tidy)
```

Now, let's generate a plot comparing the ground cover across transects.

```{r}
# this needs to be beautified

aspen_cover_types_2024_tidy %>%
  ggplot(aes(x = Transect, y = Percentage)) +
  geom_col(aes(fill = Cover_Type))
```

Let's import the data for the trees. Note that I did not import which tree was which, as I intend to add this in R.

```{r}
aspen_meadow_trees <- 
  read_excel("../../data/clean/aspen_meadow.xlsx", 
             sheet = "F24 Trees", range = "A2:M12")

View(aspen_meadow_trees)
```

```{r}
# insert t1-t4 in excel sheet, pivot longer, separate
```
