---
title: "Aspen Meadow"
output: html_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("setup.R")
```

# 2024 Ground Cover

Let's read in the Aspen Meadows ground cover data for Fall 2024.

```{r}
# reading in the ground cover data for Fall 2024
aspen_cover_2024 <- 
  read_excel("../data/clean/aspen_meadow.xlsx", 
             sheet = "F24 % Ground Cover",
             range = "A1:I11")

#view(aspen_cover_2024)
```

Next, let's select the information we need to create a plot comparing the ground cover across transects.

```{r}
# select only transect and bare ground types
aspen_cover_types_2024 <-
  aspen_cover_2024 %>%
  select(1, 3:9)

#view(aspen_cover_types_2024)
```

To more efficiently work with this graph, let's "tidy" the data.

```{r}
aspen_cover_types_2024_tidy <- aspen_cover_types_2024 %>%
  pivot_longer(cols = !Transect, names_to = "Cover_Type", 
               values_to = "Percentage")

#view(aspen_cover_types_2024_tidy)
```

Now, let's generate a plot comparing the ground cover across transects.

```{r}
aspen_cover_types_2024_tidy %>%
  ggplot(aes(x = Transect, y = Percentage)) +
  geom_col(aes(fill = Cover_Type)) +
  labs(title = "Ground Cover Type Percentage in 2024",
       subtitle = "Location: Aspen Meadows, CSU Mountain Campus, CO",
       x = "Transect Number",
       y = "Percentage of Cover Type") +
  guides(fill = guide_legend(title = "Cover Type")) +
  dark_theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = seq(1, 10, by = 1)) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot",
        axis.title.x = element_text(face = "bold", size = "10"),
        axis.title.y = element_text(face = "bold", size = "10")) +
  annotate(geom = "text", x = 4.5, y = 95,
           label = "The sum of cover types for \n Transect 2 was 80%",
           size = 2.75) +
  geom_curve(aes(x = 3, xend = 2, y = 93, yend = 81),
             arrow = arrow(length = unit(0.2, "cm")),
             size = 0.5,
             curvature = 0.4)
  

# this needs to be beautified
# install viridis package and change colors to viridis for color blind
# annotate missing data
# change x-axis to each transect number
# add graph title and subtitle, make sure to mention it is aspen meadows
```

# 2024 Trees

Let's import the data for the tree information by transect.

```{r}
# read in trees
```

```{r}
# generate a graph for trees

# scatterplot likely, x = distance, y = circumference, color = species, facet_grid by transect
```

# Historical Frequency

Let's look at the historical frequency over 1996-2004. This is the percentage of quadrats where a cover type appears. For example, if grass appeared in five quadrats out of the total ten, the historical frequency would be 50% for grass that year. If two out of ten quadrats had rocks, the historical frequency of those rocks for that year would be 20%.

```{r}
# read in historical frequency datasheet
aspen_freq <- read_excel("../data/clean/aspen_meadow.xlsx", 
    sheet = "Historical Frequency", range = "A2:K10")

#View(aspen_freq)
```

Let's tidy the dataset to make it easier to work with.

```{r}
aspen_freq_tidy <- aspen_freq %>%
  select(!"2000") %>% # removed bc no data
  pivot_longer(cols = !Cover_Type, names_to = "Year", values_to = "Values")

#View(aspen_freq_tidy)
```

Let's make a graph representing the data.

```{r}
# this graph is better for comparing full cover types
aspen_freq_tidy %>%
  ggplot(aes(x = Year, y = Values)) +
  geom_col(aes(fill = Cover_Type)) +
  labs(title = "Historical Ground Cover Frequency",
       subtitle = "Location: Aspen Meadow, CSU Mountain Campus, CO",
       x = "Year",
       y = "Percentage of Quadrats") +
  guides(fill = guide_legend(title = "Cover Type")) +
  dark_theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  theme(plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot", 
        axis.text.y = element_text(size = "7.5"))
```

An alternative graph:

```{r}
# this graph is better for comparing by individual cover types
# this graph needs to be adjusted
aspen_freq_tidy %>%
  group_by(Cover_Type) %>%
  ggplot(aes(x = Year, y = Values)) +
  geom_point(aes(color = Cover_Type)) +
  facet_wrap(~Cover_Type, ncol = 2) +
  geom_line(aes(color = Cover_Type), group = 1) +
  labs(title = "Historical Ground Cover Frequency",
       subtitle = "Location: Aspen Meadow, CSU Mountain Campus, CO",
       x = "Year",
       y = "Percentage of Quadrats") +
  guides(color = guide_legend(title = "Cover Type")) +
  dark_theme_gray() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot", 
        axis.text.x = element_text(angle = 20),
        axis.text.y = element_text(size = "7", angle = 0))
```

# Historical Relative Frequency

The historical relative frequency is the number of occurrences divided by the total number of occurrences. So if grass was found in 4 quadrats, and there were 20 quadrats in which some type of ground cover was found, the relative frequency is 4/20.

```{r}
# read in the relative frequency tab from the aspen meadows excel
aspen_rel_freq <- read_excel("../data/clean/aspen_meadow.xlsx", 
    sheet = "Historical Rel. Frequency", 
    range = "A2:K10")

#View(aspen_rel_freq)
```

Let's tidy the data to make it easier to graph.

```{r}
aspen_rel_freq_tidy <- aspen_rel_freq %>%
  select(!"2000") %>%
  pivot_longer(cols = 2:10, names_to = "Year", values_to = "Rel_Freq")

view(aspen_rel_freq_tidy)
```

Let's generate the graph for historical relative frequency.

```{r}
# generate graph similar to that in historical frequency section
aspen_rel_freq_tidy %>%
  ggplot(aes(x = Year, y = Rel_Freq)) +
  geom_col(aes(fill = Cover_Type)) +
  labs(title = "Historical Ground Cover Relative Frequency",
       subtitle = "Location: Aspen Meadows, CSU Mountain Campus, CO",
       y = "Relative Frequency (%)") +
  guides(fill = guide_legend(title = "Cover Type")) +
  dark_theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  theme(plot.title = element_text(face = "bold", size = "19"),
        plot.title.position = "plot", 
        axis.text.y = element_text(size = "7.5"))


# make sure to mention that there was some unaccounted for data, and how that affected calculations
```

# Historical Average Percent Cover

```{r}
# read in the historical avg % cover tab from the aspen meadows excel
aspen_avg_cover <- read_excel("../data/clean/aspen_meadow.xlsx", sheet = "Historical Avg % Cover", range = "A2:K10")

#View(aspen_avg_cover)
```

Let's tidy the data.

```{r}
aspen_avg_cover_tidy <- aspen_avg_cover %>%
  select(!"2000") %>%
  pivot_longer(cols = 2:10, names_to = "Year", values_to = "Avg_Cover") 

#view(aspen_avg_cover_tidy)
```

Let's graph the data.

```{r}
# generate graph similar to all above
aspen_avg_cover_tidy %>%
  ggplot(aes(x = Year, y = Avg_Cover)) +
  geom_col(aes(fill = Cover_Type)) +
  labs(title = "Historical Average Percent Ground Cover",
       subtitle = "Location: Aspen Meadow, CSU Mountain Campus, CO",
       caption = "Due to decimal rounding, results may not add up to exactly 100%",
       x = "Year",
       y = "Average Ground Cover (%)") +
  guides(color = guide_legend(title = "Cover Type")) +
  dark_theme_gray() +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot", 
        axis.text.x = element_text(angle = 20),
        axis.text.y = element_text(size = "7", angle = 0))

# make sure to mention that there was some unaccounted for data, and how that affected calculations

# annotate that the 2004 year did not add up to 100, only added to 80
```

# Phase Plots

I was tasked with developing a set of phase plots using the historical ground cover data from 1996-2004 and the ground cover data from 2024.
