---
title: "Burnt Chopped Wood"
author: "Katherine Frazer"
date: "2024-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("setup.R")
```

# 2024 Ground Cover

Let's read in the dataset.

```{r}
burnt_cover_2024 <- read_excel("../data/clean/burnt_chopped_wood.xlsx", 
    sheet = "F24 % Cover", 
    range = "A1:J11", 
    na = "NA")

View(burnt_cover_2024)
```

Let's filter it to only have relevant info, and then tidy the data for usage.

```{r}
burnt_cover_2024_tidy <-
  burnt_cover_2024 %>%
    select(1, 3:9) %>%
    pivot_longer(cols = !1, 
                 names_to = "CoverType", 
                 values_to = "Values")
view(burnt_cover_2024_tidy)
```

Let's make the graph.

```{r}
burnt_cover_2024_tidy %>%
  ggplot(aes(x = Transect, y = Values)) +
  geom_col(aes(fill = CoverType)) +
  labs(title = "Ground Cover Type Percentage in 2024",
       subtitle = "Location: Burnt Chopped Wood, CSU Mountain Campus, CO",
       caption = "Due to an error, there is no information for Transects 5 and 8.",
       x = "Transect Number",
       y = "Percentage of Cover Type") +
  guides(fill = guide_legend(title = "Cover Type")) +
  dark_theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = seq(1, 10, by = 1)) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot",
        axis.title.x = element_text(face = "bold", size = "10"),
        axis.title.y = element_text(face = "bold", size = "10"))
```

# 2024 Trees

# Historical Frequency

Let's read in the historical frequency datasheet.

```{r}
hist_freq <- read_excel("../data/clean/burnt_chopped_wood.xlsx",
                        sheet = "Historical Frequency", 
                        range = "A2:K10")

View(hist_freq)
```

Let's tidy the data.

```{r}
hist_freq_tidy <- hist_freq %>%
  select(!6) %>% #remove the year 2000
  pivot_longer(cols = !1, 
               names_to = "Year", 
               values_to = "Percentage")

view(hist_freq_tidy)
```

Let's graph.

```{r}
# this graph is better for comparing full cover types
hist_freq_tidy %>%
  ggplot(aes(x = Year, y = Percentage)) +
  geom_col(aes(fill = GroundCover)) +
  labs(title = "Historical Ground Cover Frequency",
       subtitle = "Location: Burnt Chopped Wood, CSU Mountain Campus, CO",
       x = "Year",
       y = "Percentage of Quadrats") +
  guides(fill = guide_legend(title = "Cover Type")) +
  dark_theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  theme(plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot", 
        axis.text.y = element_text(size = "7.5"))

# annotate, it seems there might have been an error in calculations for 2002 and 2003
```

```{r}
hist_freq_tidy %>%
  group_by(GroundCover) %>%
  ggplot(aes(x = Year, y = Percentage)) +
  geom_point(aes(color = GroundCover)) +
  facet_wrap(~GroundCover, ncol = 2) +
  geom_line(aes(color = GroundCover), group = 1) +
  labs(title = "Historical Ground Cover Frequency",
       subtitle = "Location: Burnt Chopped Wood, CSU Mountain Campus, CO",
       x = "Year",
       y = "Percentage of Quadrats") +
  dark_theme_gray() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot", 
        axis.text.x = element_text(angle = 20),
        axis.text.y = element_text(size = "7", angle = 0))
```

# Historical Relative Frequency

Read in the data

```{r}
his_rel_freq <- read_excel("../data/clean/burnt_chopped_wood.xlsx", 
    sheet = "Historical Rel Frequency", range = "A2:K10")

View(his_rel_freq)
```

Tidy data:

```{r}
his_rel_freq_tidy <- his_rel_freq %>%
  select(!"2000") %>%
  pivot_longer(!GroundCover, 
               names_to = "Year", 
               values_to = "Percentage")

view(his_rel_freq_tidy)
```

Graph:

```{r}
# generate graph similar to that in historical frequency section
his_rel_freq_tidy %>%
  ggplot(aes(x = Year, y = Percentage)) +
  geom_col(aes(fill = GroundCover)) +
  labs(title = "Historical Ground Cover Relative Frequency",
       subtitle = "Location: Burnt Chopped Wood, CSU Mountain Campus, CO",
       y = "Relative Frequency (%)") +
  guides(fill = guide_legend(title = "Cover Type")) +
  dark_theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  theme(plot.title = element_text(face = "bold", size = "19"),
        plot.title.position = "plot", 
        axis.text.y = element_text(size = "7.5"))


# make sure to mention that there was some unaccounted for data, and how that affected calculations (the missing transects)
```

# Historical Average Percent Cover

Data:

```{r}
hist_avg_cover <- read_excel("../data/clean/burnt_chopped_wood.xlsx", 
                             sheet = "Historical Avg % Cover", 
                             range = "A2:K10")
View(hist_avg_cover)
```

Tidy:

```{r}
hist_avg_cover_tidy <- hist_avg_cover %>%
  select(!"2000") %>%
  pivot_longer(!GroundCover, 
               names_to = "Year",
               values_to = "Percentage")

view(hist_avg_cover_tidy)
```

Graph:

```{r}
# generate graph similar to all above
hist_avg_cover_tidy %>%
  ggplot(aes(x = Year, y = Percentage)) +
  geom_col(aes(fill = GroundCover)) +
  labs(title = "Historical Average Percent Ground Cover",
       subtitle = "Location: Burnt Chopped Wood, CSU Mountain Campus, CO",
       x = "Year",
       y = "Average Ground Cover (%)") +
  guides(fill = guide_legend(title = "Cover Type")) +
  dark_theme_gray() +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold", size = "20"),
        plot.title.position = "plot", 
        axis.text.x = element_text(angle = 20),
        axis.text.y = element_text(size = "7", angle = 0))
```

# Phase Plots
